---
title: "Collaborative Filtering"
author: "Eva Tao"
date: "1/31/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

There are six units in the course HUDK4051. I will use students' ratings of these units in terms of both interest and difficulty to produce individualized suggestions about what unit to attempt next. Data of difficulty and interest were collected from 44 students in the class on a Likert-scale from 1 to 5, with 1 representing "I think this topic sounds very easy"/"I have very little interest in this topic" and 5 representing "I think this topic sounds very difficult"/"I am very interested in this topic."

Read both the interest and difficulty csv files
```{r}
I1 <- read.csv("interest.csv")
D1 <- read.csv("difficulty.csv")
```

Convert data frames to matrices
```{r, echo = FALSE}
#Remove the student ids as matrices can only contain one data type. Rename the row names with the student ids.
library(dplyr) 
I2 <- select(I1, 2:7)
I2 <- as.matrix(I2)
rownames(I2) <- I1$name
D2 <- select(D1, 2:7)
D2 <- as.matrix(D2)
rownames(D2) <- D1$name

#Missing values are coded with zero instead NA. The cosine function can not take NAs, so if we use NA, we won't be able to use most of the data. However, using zeros poses a potential problem: we assume that NA suggests that the student has no interest in the topic or think it extremely easy. We should keep this limitation in mind and continue the analysis.
I2[is.na(I2)] <- 0
D2[is.na(D2)] <- 0
```

First, let's look at the interest data. Generate a user-based similarity matrix based on cosine similarity using the ratings the students gave each unit. This matrix will represent the similarity of interests between students in the class.
```{r, echo = FALSE}
#Transpose the matrix so that multiplication occurs by students (user-based) rather than units (item-based).
I2 <- t(I2)

#Then, generate the cosine similarity values for each pair of students
#Load the lsa package to access the cosine command.
library(lsa)
I.SIM <- cosine(I2) #generates the cosine similarity values as a new matrix.
diag(I.SIM) <- NA #Since each student will be most similar to themselves, remove that information on diagnols
```

Which students are most similar to me 
```{r}
my.name <- "Yiqun Tao" #Input my name as it appears in the data set
head(rownames(I.SIM[order(I.SIM[my.name,], decreasing = TRUE),]), n = 2) #This code orders the column of the matrix according to similarity and returns the top two students who's interests are most similar to mine
```

This is a basic collaborative filter! I have used information about interest across the class to generate an individual suggestion. We may suggest students with similar interests to work together.

Now create a unit-based, rather than student-based similarity matrix for difficulty. Then use the similarity matrix to provide a suggested next unit to a student who is looking for the unit that is most similar in terms of difficulty to the "prediction" unit. 
```{r}
D.SIM <- cosine(D2)
diag(D.SIM) <- NA
head(rownames(D.SIM[order(D.SIM["pred.dif",], decreasing = TRUE),]), n = 1) 
```

Finally, educational settings have important differences to purely commercial settings such as film or product suggestions. In education we want people not to just follow their interests as they may simply choose things that are easy for them so they learn very little. To reduce this possibility with the collaborative filter create a composite measure from interest and difficulty, then construct a similarity matrix using this measure. Once you have built the similarity matrix generate a suggestion for a student who has just completed the "neural" unit.
```{r}
library(tidyr)

I3 <- gather(I1,unit,interest, 2:7)
D3 <- gather(D1, name, difficulty)

C1 <- data.frame(I3$name, I3$unit, I3$interest, D3$difficulty)
names(C1) <- c("name", "unit", "interest", "difficulty")
C1 <- filter(C1, difficulty > 0)
C2 <- select(C1, "interest", "difficulty")

#Run PCA
pc <- prcomp(C2)
#Extract PC1 loadings as new measure and attach to stid & unit
C3 <- data.frame(C1$name, C1$unit, pc$x)
C4 <- select(C3, C1.name, C1.unit, PC1)
#Remove int from unit label
C4$C1.unit <- gsub(".int", "", C4$C1.unit)

# Recreate unit by student matrix
C5 <- spread(C4, C1.name, PC1)
row.names(C5) <- C5$C1.unit
C5$C1.unit <- NULL
C5 <- as.matrix(C5)
C5 <- ifelse(is.na(C5), 0, C5)
C5 <- t(C5)

# Generate cosine similarity matrix for units
C.SIM <- cosine(C5)
diag(C.SIM) <- NA

# Search for most similar unit to "neural"
head(rownames(C.SIM[order(C.SIM["neural",], decreasing = TRUE),]), n = 1)
# The unit topic "sna" is the most similar unit to "neural" in terms of both the difficulty and interest
# So we suggest the topic "sna" for a student who just finished the "neural" topic
```

